\sysfull

### 2018-04-11 paper/intro.tex
pass over the intro.  it is too long, but the content is there

Advances in processor efficiency along with the development of energy-harvesting systems has created a new category of devices that require neither a battery nor a tethered power supply~\citep{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}. These devices operate using ambient energy, such as radio frequency transmissions~\citep{rf_powered_computing_gollakota_2014}, light~\citep{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\citep{gorlatova_sigmetrics_2014}. Incorporating compute, storage, sensing, and communication hardware~\citep{wisp5,moo}, such devices are a promising technology for use in the Internet of Things~\citep{ku_cst_2016}, in-body~\citep{nadeau_naturebio_2017} and on-body~\citep{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\citep{kicksat}.

Energy-harvesting devices create unique challenges because they operate {\em intermittently} when energy is available~\citep{hicks_isca_2017,lucia_snapl_2017}. An
energy-harvesting device buffers energy in a small storage, e.g. a capacitor~\citep{gorlatova_tmc_2013},~\citep{gunduz_commag_2014} and when a threshold amount of energy accumulates, begins operating. Harvestable energy sources are low-power compared to a platform's operating level. A device operates briefly and when buffered energy is depleted, shuts down and recharges to operate again later. As an example, recharge time may be tens of seconds in radio frequency powered medical device~\cite[Fig. 3c]{nadeau_naturebio_2017}. Moreover, charge and discharge times vary by device (due to different capacitor sizes) and some may fail $\approx$10 to $\approx$100 times per second~\citep{tan_infocom_2016},~\citep{mementos},~\citep{nvp}.

@@

\textbf{Data Consistency in Intermittent Computing.} Software in an energy-harvesting system operates in the {\em intermittent execution model}~\citep{dino,lucia_snapl_2017}, with execution interrupted by failure periods. When power fails, a device loses volatile state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e., FRAM, Flash. While capturing periodic checkpoints~\citep{mementos,quickrecall} and sleep scheduling~\citep{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave volatile and non-volatile state inconsistent, leading to unrecoverable failures~\citep{dino,edb}.

There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\citep{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\citep{hicks_isca_2017,idetic,nvp,tictpl}. New architectures require hardware changes and are inapplicable to today's systems~\citep{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now). However, they have their own limitations that need to be addressed---the core one being the rigidity of the \emph{static} decomposition into tasks.

@@

\textbf{Task Decomposition of Intermittent Programs.} Recently proposed {\em
task-based} programming and execution models~\citep{alpaca,chain} advocate for
static decomposition of a program (by a programmer) into a collection of tasks.
Tasks can include arbitrary computation and, upon completion, are guaranteed to
have executed {\em atomically}, despite arbitrarily-timed power failures.
The programmer explicitly expresses task-to-task control flow. The execution
in Figure~\ref{fig:coalesce} (top) illustrates how the decomposition into
tasks affects the time the application takes to complete. Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory, to maintain consistency of program state~\citep{chain,alpaca}. Naively minimizing the number of task transitions, (i.e. the task count), \emph{at compile time} risks creating a task that requires more energy than the device can buffer in its energy storage. Such a large task would fail to complete unless the environment provides sufficient incoming energy to supplement the energy in the capacitor. To eliminate this risk, the programmer (or the compiler) must decompose the program conservatively into many small tasks, at the cost of \emph{higher task transition overhead}.

One approach to decrease the task count, without assuming a minimum level of
incoming power always available, is to predict task energy consumption and
choose a decomposition specific to the device capacitor size. However, statically predicting energy of arbitrary input-dependent code with peripheral access is a problem without a general solution. Furthermore, a static decomposition approach prevents portability across devices with different storage capacitors.

@@

\textbf{Task Coalescing.} To address this challenge we propose a \emph{dynamic} task decomposition, that accepts any static decomposition and opportunistically coalesces tasks at runtime when there is sufficient energy. With \emph{task coalescing} atomic tasks are merged into a single (larger) task while preserving atomicity of the merged task. If power fails while a coalesced task is running, execution will restart at the first task in the group. Figure~\ref{fig:coalesce} (bottom) illustrates how dynamic coalescing changes the earlier execution and highlights the change in memory model necessary to not break atomicity.

Unfortunately, naively merging two atomic tasks might produce a non-atomic task, that may leave non-volatile memory inconsistent after a partial execution. A merge breaks atomicity when atomic merged tasks form a \emph{write-after-read (WAR)} dependency---for instance if two tasks \texttt{\{x=y+1\}; \{y=x;\}} are merged and if the power failure occurs after \texttt{x=y}, the value \texttt{x} will be increased twice when the merged task is restarted, that leads to an inconsistency.
%
%variables in non-volatile memory at \emph{run-time} that can
%break the atomicity of the \emph{virtual task}. Consider the
%example depicted in Fig.~\ref{fig:virtualization}: Tasks 1,
%2 and 3 being executed consecutively. All tasks in this
%example are atomic since they do not have WAR dependency on
%the persistent variables they are accessing, e.g. \emph{x}
%is only read and \emph{y} is only written within Task 1.
%Now, suppose three tasks have been virtualized into a single
%one at run-time, namely Task 4: since \emph{x} is now first
%\emph{read} and then \emph{written}, a WAR dependency on
%\emph{x} is introduced dynamically at
%run-time---unfortunately Task 4 is \emph{no more atomic}
%since its re-execution will not always produce the same
%results.
%
Although WAR dependencies introduced by the \emph{static} task division can be eliminated at compile time~\citep{alpaca}, those introduced \emph{dynamically} at
run-time cannot. To maintain memory consistency in merged tasks, we propose a
mechanism for \emph{memory virtualization} that buffers the updates to non-volatile variables in volatile memory before committing them at the dynamic task boundary.
%
%Therefore, we require a \emph{new execution
%model} that keeps each virtual task atomic by committing the
%modified persistent variables at the boundary defined
%dynamically at run-time---keeping the non-volatile memory
%unmodified upon a power-interrupt and preserving its
%consistency.

\textbf{Task Splitting}: The static task execution model assumes that a single task execution is guaranteed by the hardware. If this assumption is violated, due to hardware inconsistency or inaccurate task devision startegy, the system enters a non-termination loop of failed attempts to reach the next task boundery. logically, better than relying on a rondom process, energy shots arrival, to break out of this loop, the system should execute an "emergency-like" routine to terminate it and preserve forward progress.

Task coalescing removes the burden from the programmer, because it accepts any task decomposition and improves it dynamically. To further reduce the programmer effort, we propose a compiler pass for \emph{automatic decomposition} of programs into (small) \emph{atomic} tasks. The compiler identifies non-volatile variables shared across tasks as tasks are created, and instruments reads and writes of those variables, using memory virtualization, to keep the data consistent in the presence of power loss. Despite being limited to a subset of the C language, the automatic task decomposition allowed us to port several
applications to an intermittent platform with a moderate effort.

% <<<<<<< HEAD
% We develop \sys%
% \footnote{To be released at \url{http://anonymized.link}}
% %
% : a task coallescing system for task-based
% programming models on intermittently-powered devices. The
% capabilities required for effective coallescing mark our
% contributions:
% >>>>>>> 96e5d9f9b4e97ce8700c375bd8e11708870f5aba
% =======
\textbf{Contributions.} We develop \sys\footnote{To be released at \url{http://anonymized.link}}: the first task coalescing system for task-based programming models on intermittently-powered devices. The capabilities required for effective coalescing mark our contributions:
% >>>>>>> 84d7711b6c6f6424cdad97efc8aad96e52f316ff

\begin{itemize}
	\item Two adaptive coalescing strategies for selecting boundaries to coalesce based on past execution behaviour;
	\item Software memory virtualisation for transiently-powered devices that preserves atomicity of coalesced tasks;
	\item Task splitting startegy that preserve forward progress and memory consistency;
\end{itemize}

@@

We evaluated \sys end-to-end on a real energy-harvesting device~\citep{wisp} running existing benchmarks ~\citep{chain} and new programs. A comparison to a state-of-the-art task-based system~\citep{chain} showed that \sys recovers
performance on conservative task decompositions missed by systems with static decomposition.
%
%and can flexibly target a variety of platforms without rewriting code.
%
%\todo{Provide concrete numeric results supporting this claim}{Amjad/Przemek}

>>>>

Advances in processor efficiency along with the development of
energy-harvesting systems has created a new category of devices that require
neither a battery nor a tethered power
supply~\citep{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}. These
devices operate using ambient energy, such as radio frequency
transmissions~\citep{rf_powered_computing_gollakota_2014},
light~\citep{margolies_infocom_2016,margolies_tosn_2016}, and
vibration~\citep{gorlatova_sigmetrics_2014}. Incorporating compute, storage,
sensing, and communication hardware~\citep{wisp5,moo,capybara}, such devices are a
promising technology for use in the Internet of Things~\citep{ku_cst_2016},
in-body~\citep{nadeau_naturebio_2017} and
on-body~\citep{bandodkar_electroanalysis_2015} medical systems, and
energy-harvesting nano-satellites~\citep{kicksat,capybara}.

Energy-harvesting devices create unique challenges because they operate {\em
intermittently} when energy is
available~\citep{hicks_isca_2017,lucia_snapl_2017}. An energy-harvesting device
buffers energy in a small storage capacitor~\citep{gorlatova_tmc_2013,gunduz_commag_2014} and operates when a
threshold amount of energy has accumulated. 
%
Harvestable energy sources are low-power (e.g., nW to $\mu$W) compared to a platform's operating
power level (hundreds of $\mu$W to mW). A device operates briefly until it depletes its buffered energy, after which, the device shuts
down and recharges to operate again later. As an example, recharge times may be
tens of seconds in radio frequency powered medical device~\cite[Fig.
3c]{nadeau_naturebio_2017}.  The recharge and discharge time --- which corresponds to the device's inactive and active time --- varies with the energy buffering capacitor's~\cite{capybara} and some devices fail and restart operating $\approx$10 to
$\approx$100 times per second~\citep{tan_infocom_2016,mementos,nvp}.

@@

\textbf{Data Consistency in Intermittent Computing.}  Software in an
energy-harvesting system operates according to the {\em intermittent execution
model}~\citep{dino,lucia_snapl_2017}, which corresponds to the
operation-failure-restart cycles of an energy-harvesting device. In an
intermittent software execution, an {\em operating period} proceeds for an
arbitrary duration (dictated by energy availability) before being interrupted
by a {\em failure period}. In a failure period, a device loses the volatile
state in its registers, stack, SRAM, and retains the state of any non-volatile
memory, such as FRAM or Flash. While capturing periodic
checkpoints~\citep{mementos,quickrecall} and sleep
scheduling~\citep{dewdrop,hibernus,hibernusplusplus} help preserve execution
progress, failures can leave non-volatile state incorrectly, partially updated,
and may leave checkpointed volatile state inconsistent with non-volatile state.
These inconsistencies cause intermittent executio to deviate from
continuously-powered behavior, often leading to an unrecoverable
application failure~\citep{dino,edb}.

Prior work developed two main approaches to dealing with data inconsistency for
intermittently-powered devices: (i) \emph{software-based programming and
execution models}~\citep{dino,ratchet,chain,alpaca} and (ii)
\emph{hardware-based computer architecture
support}~\citep{hicks_isca_2017,idetic,nvp}.  Complex architectural changes are
expensive to design, verify, and manufacture.  New architectures are also
inapplicable to existing systems~\citep{hicks_isca_2017,nvp}. Software
approaches are simpler and applicable to existing devices today, but prior
software approaches have a number of limitations that are the focus of the work
in this paper.  A key limitation of prior software approaches that we address in this paper is the {\em inflexibility}
of a program statically decomposed into tasks.

%\textcolor{red}{tictpl is not HW, is it? also it doesn't deal with inconsistency; this paragraph doesn't include Hibernus, so it should not include CPTL either.}

@@

%Moved intel note to related work
\textbf{Task Decomposition of Intermittent Programs.} {\em
Task-based} programming and execution models require a
programmer~\citep{alpaca,chain} or a compiler~\cite{intelehcompiler} to
statically decompose a program into a collection of {\em tasks}.  A task can
include arbitrary computation and these existing systems guarantee that each
tasks executes {\em atomically}, despite arbitrarily-timed power failures.  
%
The programmer explicitly expresses task-to-task control flow, or the compiler
converts existing program control-flow into task-to-task control flow to sequence 
the tasks that it defines.
%
Figure~\ref{fig:coalesce} (top) illustrates how a program's tasks execute and
shows how tasks can impose a run-time overhead. 
%
At each task transition, the system incurs an overhead to track and atomically
commit modifications to the non-volatile memory, to maintain consistency of
program state~\citep{chain,alpaca}.  
%
The more task transitions a program executes at run time, the more overhead.

A savvy programmer may thus create very large tasks in an attempt to
minimize overhead by minimizing the number of task transitions.  However, a
very large task may require more energy to complete than a device's fixed
hardware energy buffer can hold.  Such a large task fails to completely execute
assuming only buffered energy is available (i.e., assuming the worst case).
While some environments may have sufficient harvestable energy replenish the
capacitor during operation, enabling the task to complete, assuming cooperation
from the physical environment to avoid non-termination is a risky programming
proposition.  To eliminate this risk, existing systems require the programmer
or compiler to decompose a  program into small tasks, all of which complete
using only buffered energy.  These constraints on task sizing leave an
unsatisfying dilemma: large, efficient tasks that risk non-termination, or
small tasks that are guaranteed to complete, but incur a high task transition
and commit overhead.  In this work, we propose a third way, using a novel
technique called {\em dynamic task coalescing} that efficiently executes small
tasks, avoiding unnecessary overheads and avoiding the risk of non-termination.

%Brandon: I moved Cleancut to the related work section.  THere's too much in the intro already

@@

\textbf{Task Coalescing.} In this work, we introduce a new system called \sys, which executes a task-based
intermittent program by dynamically coalescing its tasks.
\sys accepts any static decomposition (i.e., from the programmer or from a compiler) 
and coalesces its tasks at runtime. Two consecutive, coalesced tasks execute  
with no commit or task transition overhead between them, instead performing
task-end commit actions at the end second task only.
%
When there is sufficient energy to execute both tasks, two distinct, small
tasks are effectively executed as a single large task, preserving the atomicity
of both.  If power fails during a coalesced task, execution restarts from the
{\em first} of the coalesced tasks.  Figure~\ref{fig:coalesce} (bottom)
illustrates how dynamic coalescing changes the earlier execution and highlights
the change in memory model necessary to not break atomicity.

%Brandon: This is too detailed for the intro:
%Unfortunately, naively merging two atomic tasks might produce a non-atomic
%task, that may leave non-volatile memory inconsistent after a partial
%execution. A merge breaks atomicity when atomic merged tasks form a
%\emph{write-after-read (WAR)} dependency---for instance if two tasks
%\texttt{\{x=y+1\}; \{y=x;\}} are merged and if the power failure occurs after
%\texttt{x=y}, the value \texttt{x} will be increased twice when the merged task
%is restarted, that leads to an inconsistency.

%
%variables in non-volatile memory at \emph{run-time} that can
%break the atomicity of the \emph{virtual task}. Consider the
%example depicted in Fig.~\ref{fig:virtualization}: Tasks 1,
%2 and 3 being executed consecutively. All tasks in this
%example are atomic since they do not have WAR dependency on
%the persistent variables they are accessing, e.g. \emph{x}
%is only read and \emph{y} is only written within Task 1.
%Now, suppose three tasks have been virtualized into a single
%one at run-time, namely Task 4: since \emph{x} is now first
%\emph{read} and then \emph{written}, a WAR dependency on
%\emph{x} is introduced dynamically at
%run-time---unfortunately Task 4 is \emph{no more atomic}
%since its re-execution will not always produce the same
%results.
%
%Brandon: Sentence too long.
While a compiler can use static data privatization and commit instrumentation
(i.e., redo-logging) to eliminate statically identifiable, inter-task data
dependences~\citep{alpaca}, a \emph{dynamic} dependence between two coalesced
tasks requires dynamic privatization and commit actions. 
%
To ensure consistency and respect inter-task dependences, \sys uses a novel
approach to \emph{memory virtualization} that buffers non-volatile variable
updates in volatile memory during coalesced task execution, before committing
them to non-volatile memory at the dynamic task boundary.
%
%Therefore, we require a \emph{new execution
%model} that keeps each virtual task atomic by committing the
%modified persistent variables at the boundary defined
%dynamically at run-time---keeping the non-volatile memory
%unmodified upon a power-interrupt and preserving its
%consistency.

\textbf{Task Splitting}: A static task decomposition model assumes that each
single task can execute to completion.  If the hardware energy buffer provides
inadequate energy to execute each single task to completion, a program will not
terminate~\cite{cleancut}. To avoid non-termination under adversarial 
energy conditions, \sys uses a timer-based {\em partial task commit} mechanism.
Partial commit avoids non-termination by committing the intermediate state of a long-running
task that has repeatedly failed and restarted.  Partial commit violates
task atomicity, but preserves forward progress; if a programmer knows that
task atomicity is crucial to correctness, they can disable partial commit
instead risking non-termination.

%Task coalescing removes the burden from the programmer, because it accepts any
%task decomposition and improves it dynamically. To further reduce the
%programmer effort, we propose a compiler pass for \emph{automatic
%decomposition} of programs into (small) \emph{atomic} tasks. The compiler
%identifies non-volatile variables shared across tasks as tasks are created, and
%instruments reads and writes of those variables, using memory virtualization,
%to keep the data consistent in the presence of power loss.
%
%\textcolor{red}{Despite being limited to a subset of the C language, the
%automatic task decomposition allowed us to port several	applications to an
%intermittent platform with a moderate effort.}

\textbf{Contributions.} We develop \sys\footnote{To be released at
\url{http://anonymized.link}}: the first task coalescing system for task-based
programming models on intermittently-powered devices. The capabilities required
for effective coalescing mark our contributions:

\begin{itemize}
	\item \textcolor{red}{XXX} adaptive coalescing strategies for selecting boundaries to coalesce based on past execution behaviour;
	\item Software memory virtualisation for transiently-powered devices that preserves atomicity of coalesced tasks;
	\item Task splitting startegy that preserve forward progress and memory consistency;
\end{itemize}

@@

We evaluated \sys end-to-end on a real energy-harvesting device~\citep{wisp}
running existing benchmarks~\citep{chain,alpaca} and new programs. A comparison
to a state-of-the-art task-based system~\citep{chain,alpaca} showed that \sys
recovers performance on conservative task decompositions missed by systems with
static decomposition.
%
%and can flexibly target a variety of platforms without rewriting code.
%
%\todo{Provide concrete numeric results supporting this claim}{Amjad/Przemek}
